The impact of cosmic variance on simulating weak lensing surveys

### 2014-12-01 Writeup/paper_v3.tex
highlight only the newest stuff

\newcommand{\rachel}[1]{{\textcolor{red}{#1}}}
%\newcommand{\rachel}[1]{}
\newcommand{\arun}[1]{{\textcolor{blue}{#1}}}
%\newcommand{\arun}[1]{}
\newcommand{\claire}[1]{{\textcolor{magenta}{#1}}}
\def\a{\alpha}
\def\reff@jnl#1{{\rm#1\/}}

@@

\begin{abstract}
Upcoming weak lensing surveys will survey large cosmological volumes to measure the growth of cosmological structure with time and
thereby constrain dark energy.  One major systematic uncertainty in
this process is the calibration of the weak lensing shape distortions,
or shears.  Most upcoming surveys plan to test \rachel{several aspects of} their shear
estimation algorithms using sophisticated image simulations that
include 
realistic galaxy populations based on high-resolution data from the
Hubble Space Telescope ({\em HST}).  However, existing datasets
from the {\em HST} cover very small cosmological volumes, so cosmic variance could cause the galaxy populations in them
to be atypical. A narrow redshift slice from such surveys could be dominated by a single large overdensity or underdensity.  In that case, the
morphology-density relation could alter the local galaxy
populations and yield an incorrect calibration of shear estimates as a function of redshift. We directly test this scenario using the COSMOS survey, the largest-area {\em HST} survey to date, and show how the
statistical distributions of galaxy shapes and
morphological parameters (e.g., S\'{e}rsic $n$)
are influenced by redshift-dependent cosmic variance. 
% We must distinguish between environment effects from large-scale structure, which we do not wish to include in simulations, and general trends in the galaxy population with redshift. 
The typical variation in RMS ellipticity due to environmental effects
is 5 per cent \rachel{(absolute, not relative)} for redshift bins of width $\Delta z=0.05$, which could result in uncertain shear calibration at the 1 per cent level. We conclude that the cosmic variance effects are large enough to exceed the systematic error budget of future surveys, but can be mitigated with careful choice of training dataset and sufficiently large redshift binning.
\end{abstract}

@@

For the upcoming surveys to achieve their promise, their systematic error budgets must be below their statistical error budgets.
Systematic error budgets for weak lensing surveys typically include
astrophysical effects, such as intrinsic alignments of galaxy shapes
with large scale density fields \citep[e.g.,][]{2014arXiv1407.6990T} and the effect of baryons on the
matter power spectrum \rachel{\citep[e.g.,][]{2011MNRAS.415.3649V,2011MNRAS.417.2020S}}, as well as observational uncertainties such as
the ability to robustly infer shears from observed galaxy shapes or
photometric redshifts from their observed colors.
Given the expected sub-per cent statistical errors on upcoming surveys, systematic errors must be reduced from
 their typical level in the current state-of-the-art
 measurements that typically achieve $\sim 5$ per cent statistical
 errors at best
 \citep[e.g.,][]{2010A&A...516A..63S,2013MNRAS.432.2433H,2013ApJ...765...74J,2013MNRAS.432.1544M}.

 \rachel{Some types of information about and tests of shear estimation
   for ongoing and future weak lensing surveys will rely on data from
   the Hubble Space Telescope ({\em HST}).  For example, data from
   {\em HST} can be used to derive basic statistics of galaxy light
   profiles, such as joint distributions of size and the morphology.
   It can be used to infer the intrinsic distribution of galaxy
   shapes, which enters the shear estimation process either implicitly
 or explicitly depending on the method used for shear estimation
 \citep[see, e.g.,][]{2014MNRAS.439.1909V}.
 Going beyond basic information about the light profile, {\em HST} can
be used to quantify the detailed morphology of galaxies, due to its
higher resolution compared to any current or planned weak lensing
survey.  Finally, it can be used to address systematics due to colour
gradients within galaxies, which are particularly problematic when
combined with a wavelength-dependent diffraction-limited PSF
\citep[e.g.,][]{2012MNRAS.421.1385V,2013MNRAS.432.2385S}.}

@@

%The structure of the paper is as follows. In \S\ref{S:data}, we briefly describe the COSMOS catalog that we use for our analysis.
The paper is structured as follows: in Sec.~\ref{S:data}, we describe
the data that we use for this study.  In Sec.~\ref{S:methods}, we
describe our methods for deriving the relevant galaxy properties
like environment, morphology, and shape.  
Using these ingredients, we present our results in Sec.~\ref{S:results}
and discuss their implications \rachel{in Sec.~\ref{S:implications},
concluding in} Sec.~\ref{S:summary}.
\section{Data}
\label{S:data}
%In this analysis, we consider the COSMOS sample of galaxies. 
The COSMOS survey \citep{COSMOS_overview, COSMOS_generic, COSMOS_Alexie} is a flux-limited, narrow deep field survey covering a contiguous area of $1.64 \text{ deg}^2$ of sky, with images taken using the Advanced Camera for Surveys (ACS) Wide Field Channel (WFC)
in the Hubble Space Telescope (HST).  We use the COSMOS survey to
define a parent sample of galaxy images to be used for making image
simulations, following the approach taken in
\cite{2012MNRAS.420.1518M,great3}.

@@

We thus identify the regions $z=0.30-0.40$, $0.65-0.75$, and
$0.80-0.85$ as overdense; $z=0.55-0.65$ and $0.75-0.80$ as underdense;
\rachel{and we defer classication of the somewhat ambiguous range from
  $z=0.40-0.55$ until Sec.~\ref{sub:volumelimiting}. }
%Although, the $z=0.40-0.55$ bin appears to be underdense, for reasons mentioned in \S\ref{sub:volumelimiting}, we will not call them underdense.

@@

\rachel{Finally, we comment on whether the trends we
  have found could be caused by measurement noise rather than
  variations in the intrinsic shape distributions.  Measurement error
  tends to cause an increase in the RMS ellipticity due to the
  broadening of the measured ellipticity distributions.  However, this
is one of the reasons we restricted the sample to a magnitude of
23.5.  In this case, the $S/N$ of the flux measured in the galaxy
images is typically $\ge 50$.  If we consider the worst case, i.e.,
assume all galaxies have $S/N=50$ and use the Gaussian approximation
from \cite{BJ02}, the expected measurement error on the distortion is
$0.08$.  If we add this in quadrature with an RMS of $\sim 0.33$, then
the new RMS becomes $0.34$.  However, we see variations in the RMS
distortion in different environments that are as large as $0.05$, or
five times as large as this worst-case scenario due to measurement
noise.   Moreover, if measurement error were a significant issue, we
would expect it to affect the comparison of (for example) low- and
high-redshift samples in overdense regions.  In a volume-limited
sample, those will have different flux distributions and therefore
different SNR distributions.  However, there is little trend in the
RMS distortion with redshift for overdense regions, which suggests
empirically that measurement error is not a significant factor in our results.}

@@

Using the redshift distributions estimated earlier in this work (cf. Figs.~\ref{fig:redshift_fluxlimited}
and~\ref{fig:redshift_vollimited}), we obtain the overdensity estimates for
the wider bins, $\delta_{g,\text{1D}}$.
%that are shown in Figs.~\ref{fig:redshift_fluxlimited_wide} and ~\ref{fig:redshift_vollimited_wide}. 
Although the $\left[ 0.7 - 0.85\right]$ bin seems to be overdense in
the top panel of Fig.~\ref{fig:redshift_wide}, it appears to be
environmentally neutral when we \rachel{volume-limit the sample 
using the methods from Sec.~\ref{sub:volumelimiting}.}  
%(\rachel{how?  Should say here and in caption whether this is \s1, \s2, or \s3}).
Since the latter is what we use to study the galaxy morphology, we
classify $\left[ 0.7 - 0.85\right]$ bin as `neutral'.
What is surprising is the fact that the lowest and middle redshift
slices still qualify as substantial overdensities and underdensities
despite our use of $\Delta z=0.2$.

@@

\rachel{As a test of the second option, we consider whether we can use some division of the
galaxies by color or $B/T$ (as a proxy for color) in order to
determine the structural parameters of galaxies as a function of
redshift for each of them separately.  If the effects we are seeing
can be cleanly described in terms of overdense (underdense) regions
being dominated by bulge- (disk-)dominated galaxies, then division
into samples based on $B/T$ may help reduce the effects that we have
described in previous subsections.  We test this idea by dividing the
sample into bulge- and disk-dominated galaxies based on requiring
$B/T\ge $ or $<0.4$.  The resulting bulge fraction $f_\text{bulge}$,
shown as a function of redshift in Fig.~\ref{fig:fbulge_z}, 
shows a mild effect from the large-scale structures in the COSMOS field.
}
\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{fbulge_z} 
 \caption{\rachel{The fraction of galaxies in \s1 (open symbols) and
     \s3 (solid symbols) that are classified as bulge-dominated as a
     function of redshift, including the correlation with redshift
     slices classified as overdense or underdense as indicated by the
     symbol shape and color.}
         }
 \label{fig:fbulge_z}
\end{figure}

\rachel{Fig.~\ref{fig:rmsellip_bulge_disk} shows the RMS distortion as
  a function of redshift for the full \s1 sample (as shown previously) and
  for the separate bulge- and disk-dominated samples.  As shown,
  the deviations in the values due to large-scale structure are
  clearly evident in all three cases, though slightly more prominent
  for the full sample and less so for the bulge-dominated sample.  For
comparison with the results for the full sample shown in
Table~\ref{table:pvalues_all}, we compute the KS (AD) test $p$-values
for consistency of axis ratios between overdense and underdense
regions for bulge- and disk-dominated samples.  For \s1, these
$p$-values are 0.16 (0.08) for bulge-dominated galaxies and 0.005
(0.001) for disk-dominated galaxies.  For \s3, these $p$-values are
0.03 (0.03) and $10^{-4}$ ($3\times 10^{-4}$), respectively.  With the
possible exception of bulge-dominated galaxies, it seems that
separation into two morphological samples is not enough to remove the
effect of the local environment on the ensemble properties like the
intrinsic ellipticity distribution.   Due to the limited sample size,
we did not explore the possibility of an even finer morphological
division of the sample, but with deeper {\em HST} datasets, this may
be a promising path to pursue.  However, if the nature of the bulges
or disks themselves is altered by the local environment, then division
purely into bulge- or disk-dominated sample will not be helpful in
mitigating environmental effects.
}
\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{rms_ellip_momentbased_S1_bulgedisk} 
 \caption{\rachel{The RMS distortion as a function of redshift for the
     full volume-limited galaxy sample \s3 (partially-full symbols), the
     disk-dominated galaxy sample (open symbols), and the
     bulge-dominated galaxy sample (full symbols).  The fact that the
     bulge-dominated sample has a lower RMS distortion than the
     disk-dominated sample is consistent with expectations.}
         }
 \label{fig:rmsellip_bulge_disk}
\end{figure}

@@

\rachel{As mentioned previously, {\em HST} data will be used by
  current and future weak
lensing surveys to characterize the galaxy population in several
ways.  In this section, we estimate how the results shown in
Sec.~\ref{S:results} can affect estimates of shear calibration when
using {\em HST} to characterize the galaxy population.  We also
discuss the situations in which this is likely to be important for
current and future surveys from the space and ground.}

@@

\rachel{The estimates in this subsection are rough illustrations of
  the magnitudes of these effects.  Other aspects of shear calibration
  that could be affected relate to the use of {\em HST} data to
  estimate the impact of detailed galaxy morphology, or to calibrate
  the effect of colour gradients
  \citep{2012MNRAS.421.1385V,2013MNRAS.432.2385S}.  In the latter
  case, what is most relevant in this work is our finding that $B/T$
  exhibits environmental dependence, which likely translates into
  environmental dependence of colour gradients.  Unfortunately, we
  cannot directly test the strength of any colour gradient variations
  with environment using COSMOS data, due to the fact that there is
  only single-band coverage in much of its area.  We note that our
  findings may seem to be at odds with the conclusions in
  \cite{2013MNRAS.432.2385S} based on synthetic galaxy models that the
  existing area of {\em HST} coverage with $\ge 2$ bands is sufficient
  for calibration of colour
  gradients as a function of galaxy type\footnote{\rachel{\cite{2012MNRAS.421.1385V} showed that the
    fluctuations in colour gradients within the source galaxy sample
    in the Euclid weak lensing survey due to environmental effects
    will not be a significant source of uncertainty.  However, the
    question addressed in that work is different from the question
    considered here, which is the impact of fluctuations in colour
    gradients in the training sample used to derive the corrections,
    rather than in the galaxy sample to which those corrections will
    be applied.}}.  This is particularly striking given that, as shown
  there, the dominant galaxy sample with $2$-band coverage is the
  AEGIS dataset, which has substantially smaller area than COSMOS, and
  therefore should exhibit a stronger influence of cosmic variance.
  \rachel{The presence of other fields besides AEGIS (e.g., GOODS)
    should help mitigate this effect given that the large-scale
    structure in the two fields will be completely uncorrelated, but
    the number of fields is still not large and the area is dominated
    by a small number of them\footnote{\rachel{For context,
        \cite{2015APh....63...81N} 
        showed that for tests of photometric redshift quality, the
        number of spectroscopic fields of this size that would be
        required for future surveys to avoid the influence of cosmic
        variance ranges from many tens to hundreds}}.}  It seems that a future study using
  an {\em HST} field with more than one band would be warranted, to
  test whether the discrepancy between our findings and those of
  \cite{2013MNRAS.432.2385S} arises from a weak connection between
  colour gradients and $B/T$ (meaning that our results do not imply a
  problem with colour gradient calibration), or from some unrealistic
  aspect to the synthetic models used in that work.}

\subsection{\rachel{Effective impact on current and future surveys}}

\rachel{In addition to the order of magnitude of these effects
  presented in the previous subsection, it is important to bear in
  mind how current and future surveys plan to use {\em HST} data.}

@@

\rachel{However, once the Euclid survey is
  underway, the derivation of simple galaxy morphology and the
  intrinsic ellipticity distribution will be based on the 40~deg$^2$
  Euclid deep field, which goes two magnitudes fainter than the rest of
  the survey.  With that data in hand, the Euclid weak lensing
  results will be less reliant on the much smaller and more
  cosmic variance-limited {\em HST} fields, using them only for colour
  gradient calibration and estimates of the impact of detailed galaxy
  morphology (since the {\em HST} resolution is higher than that
  of Euclid).  The findings of \cite{2014MNRAS.439.1909V} demonstrate that the area of
  this deep field is sufficient to accomplish the goal of determining
  the intrinsic ellipticity distribution at the accuracy required for
  shear calibration purposes for the Euclid survey.}

\rachel{In contrast, for ground-based surveys such
  as LSST, high-resolution space data will play a more important role
  in the understanding of the galaxy 
  intrinsic ellipticity and morphology distributions, since even a deep
  field in a ground-based survey faces fundamental resolution limits
  that prevent the derivation of detailed information about the faint
  galaxy population.  Galaxies that are near the resolution limit for
  a ground-based survey are still very well-resolved in {\em HST},
  making it the best resource for detailed information about them.
  Once the Euclid deep field is publicly available, the information
  from it will be beneficial to ground-based surveys as well.}

\rachel{Finally, our results for the effect on the intrinsic
  ellipticity distributions suggest that even current surveys should
  be careful to avoid this effect.  While use of COSMOS without
  accounting for the effect that we have identified will cause a bias
  that is similar to the final requirements on the shear systematic
  errors for Stage III surveys, there are other elements in the
  systematic error budget, so ideally this effect should be mitigated
  somewhat using, for example, the wider redshift binning strategy
  that we tested in Sec.~\ref{sec:mitigation}.  Use of significantly smaller fields
  than COSMOS, while possibly helpful in building up a deeper galaxy
  sample, should naturally increase the impact of cosmic
  variance on the training dataset, which should be quantified as part
  of the systematic error budget.}

@@

It is important to keep in mind the nature of COSMOS with respect to
other possible {\em HST} training samples.  COSMOS represents the
largest contiguous field surveyed by the {\em HST}, with the sizes of
other major {\em HST} fields such as GOODS and AEGIS lagging by at least a
\rachel{factor of 9 (for AEGIS, and more for GOODS).  The GEMS field
  comes the closest to that of COSMOS, but is still a factor of 6
  smaller and may be less useful for determination of colour gradients
  given that one of the bands, F850LP, is relatively shallow.}.
Hence, if cosmic variance due to structures along 
the line-of-sight in COSMOS are problematic for its use as a training
sample for weak lensing simulations, studies that use even smaller
area training samples are even more prone to errors, with the UDF
serving as an extreme case.  If the size of the {\em HST} survey is
small enough, there is no reason {\em a priori} to suppose that the
galaxy population is typical even when using all galaxies along the
line of sight, without any division into redshift bins.  Of course,
future surveys are unlikely to pick just a single survey to serve as
the basis for their image simulation training sample, but rather
will combine as many as possible.  Combining multiple surveys will
reduce the cosmic variance and therefore the significance of the
effects discussed in this work.
However, as COSMOS is significantly larger than other \emph{HST} surveys, combining COSMOS with smaller fields is unlikely to ameliorate this effect.
Thus, it will be important to carefully choose the size of redshift slices used to
derive properties of the galaxy population so as to be minimally
affected by this issue.

>>>>

%\newcommand{\rachel}[1]{{\textcolor{red}{#1}}}
\newcommand{\rachel}[1]{}
\newcommand{\rachelnew}[1]{{\textcolor{red}{#1}}}
\newcommand{\arun}[1]{{\textcolor{blue}{#1}}}
%\newcommand{\arun}[1]{}
\newcommand{\claire}[1]{{\textcolor{magenta}{#1}}}
\def\a{\alpha}
\def\reff@jnl#1{{\rm#1\/}}

@@

\begin{abstract}
Upcoming weak lensing surveys will survey large cosmological volumes to measure the growth of cosmological structure with time and
thereby constrain dark energy.  One major systematic uncertainty in
this process is the calibration of the weak lensing shape distortions,
or shears.  Most upcoming surveys plan to test \rachel{several aspects of} their shear
estimation algorithms using sophisticated image simulations that
include 
realistic galaxy populations based on high-resolution data from the
Hubble Space Telescope ({\em HST}).  However, existing datasets
from the {\em HST} cover very small cosmological volumes, so cosmic variance could cause the galaxy populations in them
to be atypical. A narrow redshift slice from such surveys could be dominated by a single large overdensity or underdensity.  In that case, the
morphology-density relation could alter the local galaxy
populations and yield an incorrect calibration of shear estimates as a function of redshift. We directly test this scenario using the COSMOS survey, the largest-area {\em HST} survey to date, and show how the
statistical distributions of galaxy shapes and
morphological parameters (e.g., S\'{e}rsic $n$)
are influenced by redshift-dependent cosmic variance. 
% We must distinguish between environment effects from large-scale structure, which we do not wish to include in simulations, and general trends in the galaxy population with redshift. 
The typical variation in RMS ellipticity due to environmental effects
is 5 per cent \rachelnew{(absolute, not relative)} for redshift bins of width $\Delta z=0.05$, which could result in uncertain shear calibration at the 1 per cent level. We conclude that the cosmic variance effects are large enough to exceed the systematic error budget of future surveys, but can be mitigated with careful choice of training dataset and sufficiently large redshift binning.
\end{abstract}

@@

For the upcoming surveys to achieve their promise, their systematic error budgets must be below their statistical error budgets.
Systematic error budgets for weak lensing surveys typically include
astrophysical effects, such as intrinsic alignments of galaxy shapes
with large scale density fields \citep[e.g.,][]{2014arXiv1407.6990T} and the effect of baryons on the
matter power spectrum \rachelnew{\citep[e.g.,][]{2011MNRAS.415.3649V,2011MNRAS.417.2020S}}, as well as observational uncertainties such as
the ability to robustly infer shears from observed galaxy shapes or
photometric redshifts from their observed colors.
Given the expected sub-per cent statistical errors on upcoming surveys, systematic errors must be reduced from
 their typical level in the current state-of-the-art
 measurements that typically achieve $\sim 5$ per cent statistical
 errors at best
 \citep[e.g.,][]{2010A&A...516A..63S,2013MNRAS.432.2433H,2013ApJ...765...74J,2013MNRAS.432.1544M}.

 \rachel{Some types of information about and tests of shear estimation
   for ongoing and future weak lensing surveys will rely on data from
   the Hubble Space Telescope ({\em HST}).  For example, data from
   {\em HST} can be used to derive basic statistics of galaxy light
   profiles, such as joint distributions of size and the morphology.
   It can be used to infer the intrinsic distribution of galaxy
   shapes, which enters the shear estimation process either implicitly
 or explicitly depending on the method used for shear estimation
 \rachelnew{\citep[see, e.g.,][]{2014MNRAS.439.1909V}.}
 Going beyond basic information about the light profile, {\em HST} can
be used to quantify the detailed morphology of galaxies, due to its
higher resolution compared to any current or planned weak lensing
survey.  Finally, it can be used to address systematics due to colour
gradients within galaxies, which are particularly problematic when
combined with a wavelength-dependent diffraction-limited PSF
\citep[e.g.,][]{2012MNRAS.421.1385V,2013MNRAS.432.2385S}.}

@@

%The structure of the paper is as follows. In \S\ref{S:data}, we briefly describe the COSMOS catalog that we use for our analysis.
The paper is structured as follows: in Sec.~\ref{S:data}, we describe
the data that we use for this study.  In Sec.~\ref{S:methods}, we
describe our methods for deriving the relevant galaxy properties
like environment, morphology, and shape.  
Using these ingredients, we present our results in Sec.~\ref{S:results}
and discuss their implications \rachelnew{in Sec.~\ref{S:implications},
concluding in} Sec.~\ref{S:summary}.
\section{Data}
\label{S:data}
%In this analysis, we consider the COSMOS sample of galaxies. 
The COSMOS survey \citep{COSMOS_overview, COSMOS_generic, COSMOS_Alexie} is a flux-limited, narrow deep field survey covering a contiguous area of $1.64 \text{ deg}^2$ of sky, with images taken using the Advanced Camera for Surveys (ACS) Wide Field Channel (WFC)
in the Hubble Space Telescope (HST).  We use the COSMOS survey to
define a parent sample of galaxy images to be used for making image
simulations, following the approach taken in
\cite{2012MNRAS.420.1518M,great3}.

@@

We thus identify the regions $z=0.30-0.40$, $0.65-0.75$, and
$0.80-0.85$ as overdense; $z=0.55-0.65$ and $0.75-0.80$ as underdense;
\rachelnew{and we defer classication of the somewhat ambiguous range from
  $z=0.40-0.55$ until Sec.~\ref{sub:volumelimiting}. }
%Although, the $z=0.40-0.55$ bin appears to be underdense, for reasons mentioned in \S\ref{sub:volumelimiting}, we will not call them underdense.

@@

\rachelnew{Finally, we comment on whether the trends we
  have found could be caused by measurement noise rather than
  variations in the intrinsic shape distributions.  Measurement error
  tends to cause an increase in the RMS ellipticity due to the
  broadening of the measured ellipticity distributions.  However, this
is one of the reasons we restricted the sample to a magnitude of
23.5.  In this case, the $S/N$ of the flux measured in the galaxy
images is typically $\ge 50$.  If we consider the worst case, i.e.,
assume all galaxies have $S/N=50$ and use the Gaussian approximation
from \cite{BJ02}, the expected measurement error on the distortion is
$0.08$.  If we add this in quadrature with an RMS of $\sim 0.33$, then
the new RMS becomes $0.34$.  However, we see variations in the RMS
distortion in different environments that are as large as $0.05$, or
five times as large as this worst-case scenario due to measurement
noise.   Moreover, if measurement error were a significant issue, we
would expect it to affect the comparison of (for example) low- and
high-redshift samples in overdense regions.  In a volume-limited
sample, those will have different flux distributions and therefore
different SNR distributions.  However, there is little trend in the
RMS distortion with redshift for overdense regions, which suggests
empirically that measurement error is not a significant factor in our results.}

@@

Using the redshift distributions estimated earlier in this work (cf. Figs.~\ref{fig:redshift_fluxlimited}
and~\ref{fig:redshift_vollimited}), we obtain the overdensity estimates for
the wider bins, $\delta_{g,\text{1D}}$.
%that are shown in Figs.~\ref{fig:redshift_fluxlimited_wide} and ~\ref{fig:redshift_vollimited_wide}. 
Although the $\left[ 0.7 - 0.85\right]$ bin seems to be overdense in
the top panel of Fig.~\ref{fig:redshift_wide}, it appears to be
environmentally neutral when we \rachelnew{volume-limit the sample 
using the methods from Sec.~\ref{sub:volumelimiting}.}  
%(\rachel{how?  Should say here and in caption whether this is \s1, \s2, or \s3}).
Since the latter is what we use to study the galaxy morphology, we
classify $\left[ 0.7 - 0.85\right]$ bin as `neutral'.
What is surprising is the fact that the lowest and middle redshift
slices still qualify as substantial overdensities and underdensities
despite our use of $\Delta z=0.2$.

@@

\rachelnew{As a test of the second option, we consider whether we can use some division of the
galaxies by color or $B/T$ (as a proxy for color) in order to
determine the structural parameters of galaxies as a function of
redshift for each of them separately.  If the effects we are seeing
can be cleanly described in terms of overdense (underdense) regions
being dominated by bulge- (disk-)dominated galaxies, then division
into samples based on $B/T$ may help reduce the effects that we have
described in previous subsections.  We test this idea by dividing the
sample into bulge- and disk-dominated galaxies based on requiring
$B/T\ge $ or $<0.4$.  The resulting bulge fraction $f_\text{bulge}$,
shown as a function of redshift in Fig.~\ref{fig:fbulge_z}, 
shows a mild effect from the large-scale structures in the COSMOS field.
}
\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{fbulge_z} 
 \caption{\rachelnew{The fraction of galaxies in \s1 (open symbols) and
     \s3 (solid symbols) that are classified as bulge-dominated as a
     function of redshift, including the correlation with redshift
     slices classified as overdense or underdense as indicated by the
     symbol shape and color.}
         }
 \label{fig:fbulge_z}
\end{figure}

\rachelnew{Fig.~\ref{fig:rmsellip_bulge_disk} shows the RMS distortion as
  a function of redshift for the full \s1 sample (as shown previously) and
  for the separate bulge- and disk-dominated samples.  As shown,
  the deviations in the values due to large-scale structure are
  clearly evident in all three cases, though slightly more prominent
  for the full sample and less so for the bulge-dominated sample.  For
comparison with the results for the full sample shown in
Table~\ref{table:pvalues_all}, we compute the KS (AD) test $p$-values
for consistency of axis ratios between overdense and underdense
regions for bulge- and disk-dominated samples.  For \s1, these
$p$-values are 0.16 (0.08) for bulge-dominated galaxies and 0.005
(0.001) for disk-dominated galaxies.  For \s3, these $p$-values are
0.03 (0.03) and $10^{-4}$ ($3\times 10^{-4}$), respectively.  With the
possible exception of bulge-dominated galaxies, it seems that
separation into two morphological samples is not enough to remove the
effect of the local environment on the ensemble properties like the
intrinsic ellipticity distribution.   Due to the limited sample size,
we did not explore the possibility of an even finer morphological
division of the sample, but with deeper {\em HST} datasets, this may
be a promising path to pursue.  However, if the nature of the bulges
or disks themselves is altered by the local environment, then division
purely into bulge- or disk-dominated sample will not be helpful in
mitigating environmental effects.
}
\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{rms_ellip_momentbased_S1_bulgedisk} 
 \caption{\rachelnew{The RMS distortion as a function of redshift for the
     full volume-limited galaxy sample \s3 (partially-full symbols), the
     disk-dominated galaxy sample (open symbols), and the
     bulge-dominated galaxy sample (full symbols).  The fact that the
     bulge-dominated sample has a lower RMS distortion than the
     disk-dominated sample is consistent with expectations.}
         }
 \label{fig:rmsellip_bulge_disk}
\end{figure}

@@

\rachel{As mentioned previously, {\em HST} data will be used by
\rachelnew{  current and} future weak
lensing surveys to characterize the galaxy population in several
ways.  In this section, we estimate how the results shown in
Sec.~\ref{S:results} can affect estimates of shear calibration when
using {\em HST} to characterize the galaxy population.  We also
discuss the situations in which this is likely to be important for
current and future surveys from the space and ground.}

@@

\rachel{The estimates in this subsection are rough illustrations of
  the magnitudes of these effects.  Other aspects of shear calibration
  that could be affected relate to the use of {\em HST} data to
  estimate the impact of detailed galaxy morphology, or to calibrate
  the effect of colour gradients
  \citep{2012MNRAS.421.1385V,2013MNRAS.432.2385S}.  In the latter
  case, what is most relevant in this work is our finding that $B/T$
  exhibits environmental dependence, which likely translates into
  environmental dependence of colour gradients.  Unfortunately, we
  cannot directly test the strength of any colour gradient variations
  with environment using COSMOS data, due to the fact that there is
  only single-band coverage in much of its area.  We note that our
  findings may seem to be at odds with the conclusions in
  \cite{2013MNRAS.432.2385S} based on synthetic galaxy models that the
  existing area of {\em HST} coverage with $\ge 2$ bands is sufficient
  for calibration of colour
  gradients \rachelnew{ as a function of galaxy type}\footnote{\rachel{\cite{2012MNRAS.421.1385V} showed that the
    fluctuations in colour gradients within the source galaxy sample
    in the Euclid weak lensing survey due to environmental effects
    will not be a significant source of uncertainty.  However, the
    question addressed in that work is different from the question
    considered here, which is the impact of fluctuations in colour
    gradients in the training sample used to derive the corrections,
    rather than in the galaxy sample to which those corrections will
    be applied.}}.  This is particularly striking given that, as shown
  there, the dominant galaxy sample with $2$-band coverage is the
  AEGIS dataset, which has substantially smaller area than COSMOS, and
  therefore should exhibit a stronger influence of cosmic variance.
  \rachelnew{The presence of other fields besides AEGIS (e.g., GOODS)
    should help mitigate this effect given that the large-scale
    structure in the two fields will be completely uncorrelated, but
    the number of fields is still not large and the area is dominated
    by a small number of them\footnote{\rachelnew{For context,
        \cite{2015APh....63...81N} 
        showed that for tests of photometric redshift quality, the
        number of spectroscopic fields of this size that would be
        required for future surveys to avoid the influence of cosmic
        variance ranges from many tens to hundreds}}.}  It seems that a future study using
  an {\em HST} field with more than one band would be warranted, to
  test whether the discrepancy between our findings and those of
  \cite{2013MNRAS.432.2385S} arises from a weak connection between
  colour gradients and $B/T$ (meaning that our results do not imply a
  problem with colour gradient calibration), or from some unrealistic
  aspect to the synthetic models used in that work.}

\subsection{\rachel{Effective impact on \rachelnew{current and} future surveys}}

\rachel{In addition to the order of magnitude of these effects
  presented in the previous subsection, it is important to bear in
  mind how \rachelnew{current and} future surveys plan to use {\em HST} data.}

@@

\rachel{However, once the Euclid survey is
  underway, the derivation of simple galaxy morphology and the
  intrinsic ellipticity distribution will be based on the 40~deg$^2$
  Euclid deep field, which goes two magnitudes fainter than the rest of
  the survey.  With that data in hand, the Euclid weak lensing
  results will be less reliant on the much smaller and more
  cosmic variance-limited {\em HST} fields, using them only for colour
  gradient calibration and estimates of the impact of detailed galaxy
  morphology (since the {\em HST} resolution is higher than that
  of Euclid). \rachelnew{ The findings of \cite{2014MNRAS.439.1909V} demonstrate that the area of
  this deep field is sufficient to accomplish the goal of determining
  the intrinsic ellipticity distribution at the accuracy required for
  shear calibration purposes for the Euclid survey.}}

\rachel{In contrast, for ground-based surveys such
  as LSST, high-resolution space data will play a more important role
  in the understanding of the galaxy 
  intrinsic ellipticity and morphology distributions, since even a deep
  field in a ground-based survey faces fundamental resolution limits
  that prevent the derivation of detailed information about the faint
  galaxy population.  Galaxies that are near the resolution limit for
  a ground-based survey are still very well-resolved in {\em HST},
  making it the best resource for detailed information about them.}
  \rachelnew{Once the Euclid deep field is publicly available, the information
  from it will be beneficial to ground-based surveys as well.}

\rachelnew{Finally, our results for the effect on the intrinsic
  ellipticity distributions suggest that even current surveys should
  be careful to avoid this effect.  While use of COSMOS without
  accounting for the effect that we have identified will cause a bias
  that is similar to the final requirements on the shear systematic
  errors for Stage III surveys, there are other elements in the
  systematic error budget, so ideally this effect should be mitigated
  somewhat using, for example, the wider redshift binning strategy
  that we tested in Sec.~\ref{sec:mitigation}.  Use of significantly smaller fields
  than COSMOS, while possibly helpful in building up a deeper galaxy
  sample, should naturally increase the impact of cosmic
  variance on the training dataset, which should be quantified as part
  of the systematic error budget.}

@@

It is important to keep in mind the nature of COSMOS with respect to
other possible {\em HST} training samples.  COSMOS represents the
largest contiguous field surveyed by the {\em HST}, with the sizes of
other major {\em HST} fields such as GOODS and AEGIS lagging by at least a
\rachelnew{factor of 9 (for AEGIS, and more for GOODS).  The GEMS field
  comes the closest to that of COSMOS, but is still a factor of 6
  smaller and may be less useful for determination of colour gradients
  given that one of the bands, F850LP, is relatively shallow.}.
Hence, if cosmic variance due to structures along 
the line-of-sight in COSMOS are problematic for its use as a training
sample for weak lensing simulations, studies that use even smaller
area training samples are even more prone to errors, with the UDF
serving as an extreme case.  If the size of the {\em HST} survey is
small enough, there is no reason {\em a priori} to suppose that the
galaxy population is typical even when using all galaxies along the
line of sight, without any division into redshift bins.  Of course,
future surveys are unlikely to pick just a single survey to serve as
the basis for their image simulation training sample, but rather
will combine as many as possible.  Combining multiple surveys will
reduce the cosmic variance and therefore the significance of the
effects discussed in this work.
However, as COSMOS is significantly larger than other \emph{HST} surveys, combining COSMOS with smaller fields is unlikely to ameliorate this effect.
Thus, it will be important to carefully choose the size of redshift slices used to
derive properties of the galaxy population so as to be minimally
affected by this issue.

A final consideration is the question of how applicable these results
using volume-limited samples are to simulations of upcoming weak
lensing surveys, which will exclusively use flux-limited samples.  For
our analysis, the volume-limiting sample was necessary to avoid
complications due to varying galaxy populations in each redshift
slice, allowing us to isolate purely environmental effects.  In
principle, if the morphology-density correlations that we have
identified turn out to not exist for intrinsically fainter galaxy
populations, then at low redshift (where a flux limited sample will
include galaxies that are intrinsically much fainter than at high
redshift), the effects will be less serious for upcoming lensing
surveys.  However, we do not have any particular reason to believe
that these effects will vanish for fainter galaxies.  Moreover, at
higher redshift where only intrinsically bright galaxies can be seen,
the effect should be present at a level similar to what we have found
here.  Since higher redshift galaxies tend to dominate cosmological
shear estimates (due to their higher shears), our findings will be
important to take into account.  \rachelnew{It would also be advisable to
  carry out a future study of this effect at higher redshift (beyond
  $0.85$), using for example the data from the CANDELS survey.}
